# Script and session info

Script name: PathwayVisualization.Rmd

Script purpose: Pathway loading + preparing for merge

Author: Martina Kutmon

Date Created: 2020-04-08

## Session info:
R version 3.6.3 (2020-02-29)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows 10 x64 (build 17763)
Packages: rstudioapi v0.11, RCy3 v2.6.3, rWikiPathways v1.6.1, clusterProfiler v3.14.3, tidyr v1.0.2, dplyr v0.8.4, biomaRt v2.42.0

Cytoscape version 3.7.2, WikiPathways app version 3.3.7

# Install required packages
```{r}
# install.packages("BiocManager")
# BiocManager::install(c("rstudioapi","RCy3", "rWikiPathways", "clusterProfiler", "tidyr", "dplyr", "biomaRt"))
library(rstudioapi)
library(RCy3)
library(rWikiPathways)
library(clusterProfiler)
library(tidyr)
library(dplyr)
library(biomaRt)
```

# Set up environment
```{r}
#clear workspace and set string as factors to false
rm(list=ls())
options(stringsAsFactors = F)

#set working directroy
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

# Get WP pathway-gene associations
```{r}
wp.hs.gmt <- rWikiPathways::downloadPathwayArchive(organism="Homo sapiens", format = "gmt")
wp2gene <- clusterProfiler::read.gmt(wp.hs.gmt)
wp2gene <- wp2gene %>% tidyr::separate(ont, c("name","version","wpid","org"), "%")
wpid2gene <- wp2gene %>% dplyr::select(wpid,gene) #TERM2GENE
wpid2name <- wp2gene %>% dplyr::select(wpid,name) #TERM2NAME
```

# Filter pathways of interest
```{r}
pathways <- c("WP45","WP179", "WP707")
selected_wpid2gene <- wpid2gene[wpid2gene$wpid %in% pathways,]
colnames(selected_wpid2gene) <- c("source","target")
```

# Create gene-pathway network
```{r}
RCy3::cytoscapePing()
suid <- RCy3::createNetworkFromDataFrames(edges=selected_wpid2gene, title="gene-pathway network", collection = "gene-pathway network")
wpid2name['Type'] <- "pathway"
RCy3::loadTableData(wpid2name, data.key.column = "wpid", table.key.column = "id")

ensembl <- useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl", mirror = "useast")
genes <- getBM(attributes = c('ensembl_gene_id', 'hgnc_symbol', 'entrezgene_id'), filters = 'entrezgene_id', values = selected_wpid2gene$target, mart = ensembl)
colnames(genes) <- c("Ensembl","name","id")
RCy3::loadTableData(genes, data.key.column = "id", table.key.column = "id")

RCy3::setNodeColorMapping("Type", table.column.values = "pathway", colors="#f7b5b8", mapping.type="d", default.color="#a1c8ed")
RCy3::setNodeShapeDefault('ELLIPSE')
RCy3::setNodeSizeDefault(30)

RCy3::toggleGraphicsDetails()
```